Weighted Model Counting with Conditional Measures

### 2020-07-28 paper.tex
A bugfix (and a few other things)

\begin{definition}[Special functions]
  \phantom{}
  \begin{itemize}
  \item unit $1\colon 2^\emptyset \to \mathbb{R}_{\ge 0}$, $1(\emptyset) = 1$.
  \item zero $0\colon 2^\emptyset \to \mathbb{R}_{\ge 0}$, $0(\emptyset) = 0$.
  \item constant $[a]\colon 2^{\{a\}} \to \mathbb{R}_{\ge 0}$, $[a](\emptyset) =
    0$, $[a](\{a\}) = 1$.
  \end{itemize}
\end{definition}

@@

\begin{theorem}
  A measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is factorable if and only
  if
  \begin{equation} \label{eq:wmccondition}
  M(u \land v) = M(u)M(v)
  \end{equation}
  for all distinct $u, v \in U \cup \{ \neg w \mid w \in U \}$ such that $u \ne
  \neg v$.
\end{theorem}
\begin{proof}
  \todo[inline]{Check if the proof is valid for zero weights.}
  ($\Leftarrow$) For each $x \in U$, let $W_x\colon 2^{\{x\}} \to
  \mathbb{R}_{\ge 0}$ be defined by $W_x(\{ x \}) = M(x)$ and $W_x(\emptyset) =
  M(\neg x)$. Let $M_W$ be the measure induced by
  \[
    W = \prod_{x \in U} W_x.
  \]
  We will show that $M = M_W$. First, note that $M_w(\bot) = 0 = M(\bot)$ by
  \cref{def:wmc,prop:measure}. Second, let
  \[
    a = \bigwedge_{u \in U} a_u
  \]
  be an atom in $2^{2^U}$ such that $a_u \in \{ u, \neg u \}$ for all $u \in U$.
  Then
  \[
    M_W(a) = W(\{ a_u \mid u \in U \}) = \prod_{u \in U} W_u(\{a_u\}) = \prod_{u
      \in U} M(a_u) = M \left(\bigwedge_{u \in U} a_u \right) = M(a)
  \]

@@

  ($\Rightarrow$) For the other direction, we are given a factored weight
  function
  \[
    W = \prod_{x \in U} W_x,
  \]
  and we want to show that its induced measure $M_W$ satisfies
  \cref{eq:wmccondition}. Let $k_u, k_v \in U \cup \{ \neg w \mid w \in U \}$ be
  such that $k_u \in \{ u, \neg u \}$, $k_v \in \{ v, \neg v \}$, and $u \ne v$.
  We then want to show that
  \begin{equation} \label{eq:to_prove3}
    M_W(k_u \land k_v) = M_W(k_u)M_W(k_v)
  \end{equation}
  which is equivalent to
  \begin{equation} \label{eq:to_prove4}
    M_W(k_u \land k_v) \cdot M_W(\neg k_u \land \neg k_v) = M_W(k_u \land \neg k_v) \cdot M_w(\neg k_u \land k_v)
  \end{equation}
  by \cref{lemma:before_theorem}. Then
  \begin{align*}
    M_W(k_u \land k_v) &= \sum_{\{a\} \le k_u \land k_v} W(a) = \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U} W_x(a) \\
                        &= \sum_{\{a\} \le k_u \land k_v} W_u(a_u)W_v(a_v) \prod_{x \in U \setminus \{ u, v \}} W_x(a) = \sum_{\{a\} \le k_u \land k_v} W_u(k_u)W_v(k_v) \prod_{x \in U \setminus \{ u, v \}} W_x(a) \\
    &= W_u(k_u)W_v(k_v) \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U \setminus \{ u, v \}} W_x(a) = W_u(k_u)W_v(k_v)C,
  \end{align*}
  where $C$ denotes the part of $M_W(k_u \land k_v)$ that will be the same for
  $M_W(\neg k_u \land k_v)$, $M_W(k_u \land \neg k_v)$, and $M_W(\neg k_u \land
  \neg k_v)$ as well. But then \cref{eq:to_prove4} becomes
  \[
    W_u(k_u)W_v(k_v)W_u(\neg k_u)W_v(\neg k_v)C^2 = W_u(k_u)W_v(\neg
    k_v)W_u(\neg k_u)W_v(k_v)C^2
  \]
  which is trivially true.
\end{proof}

@@

\subsection{Adjusting ADDMC to Work with the New Format}

@@

>>>>

\begin{definition}[Special functions]
  \phantom{}
  \begin{itemize}
  \item unit $1\colon 2^\emptyset \to \mathbb{R}_{\ge 0}$, $1(\emptyset) = 1$.
  \item zero $0\colon 2^\emptyset \to \mathbb{R}_{\ge 0}$, $0(\emptyset) = 0$.
  \item logical constant $[a]\colon 2^{\{a\}} \to \mathbb{R}_{\ge 0}$,
    $[a](\emptyset) = 0$, $[a](\{a\}) = 1$.
  \end{itemize}
\end{definition}

@@

\begin{theorem}
  A measure $M\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is factorable if and only
  if
  \begin{equation} \label{eq:wmccondition}
  M(u \land v) = M(u)M(v)
  \end{equation}
  for all distinct $u, v \in U \cup \{ \neg w \mid w \in U \}$ such that $u \ne
  \neg v$.
\end{theorem}
\begin{proof}
  ($\Leftarrow$) For each $x \in U$, let $W_x\colon 2^{\{x\}} \to
  \mathbb{R}_{\ge 0}$ be defined by $W_x(\{ x \}) = M(x)$ and $W_x(\emptyset) =
  M(\neg x)$. Let $M_W$ be the measure induced by
  \[
    W = \prod_{x \in U} W_x.
  \]
  We will show that $M = M_W$. First, note that $M_w(\bot) = 0 = M(\bot)$ by
  \cref{def:wmc,prop:measure}. Second, let
  \[
    a = \bigwedge_{u \in U} a_u
  \]
  be an atom in $2^{2^U}$ such that $a_u \in \{ u, \neg u \}$ for all $u \in U$.
  Then
  \[
    M_W(a) = W(\{ a_u \mid u \in U \}) = \prod_{u \in U} W_u(\{a_u\}) = \prod_{u
      \in U} M(a_u) = M \left(\bigwedge_{u \in U} a_u \right) = M(a)
  \]

@@

  ($\Rightarrow$) For the other direction, we are given a factored weight
  function
  \[
    W = \prod_{x \in U} W_x,
  \]
  and we want to show that its induced measure $M_W$ satisfies
  \cref{eq:wmccondition}. Let $k_u, k_v \in U \cup \{ \neg w \mid w \in U \}$ be
  such that $k_u \in \{ u, \neg u \}$, $k_v \in \{ v, \neg v \}$, and $u \ne v$.
  We then want to show that
  \begin{equation} \label{eq:to_prove3}
    M_W(k_u \land k_v) = M_W(k_u)M_W(k_v)
  \end{equation}
  which is equivalent to
  \begin{equation} \label{eq:to_prove4}
    M_W(k_u \land k_v) \cdot M_W(\neg k_u \land \neg k_v) = M_W(k_u \land \neg k_v) \cdot M_w(\neg k_u \land k_v)
  \end{equation}
  by \cref{lemma:before_theorem}. Then
  \begin{align*}
    M_W(k_u \land k_v) &= \sum_{\{a\} \le k_u \land k_v} W(a) = \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U} W_x(a) \\
                        &= \sum_{\{a\} \le k_u \land k_v} W_u(a_u)W_v(a_v) \prod_{x \in U \setminus \{ u, v \}} W_x(a) = \sum_{\{a\} \le k_u \land k_v} W_u(k_u)W_v(k_v) \prod_{x \in U \setminus \{ u, v \}} W_x(a) \\
    &= W_u(k_u)W_v(k_v) \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U \setminus \{ u, v \}} W_x(a) = W_u(k_u)W_v(k_v)C,
  \end{align*}
  where
  \[
    C = \sum_{\{a\} \le k_u \land k_v} \prod_{x \in U \setminus \{ u, v \}} W_x(a),
  \]
  and is the same for $M_W(\neg k_u \land k_v)$, $M_W(k_u \land \neg k_v)$, and
  $M_W(\neg k_u \land \neg k_v)$ as well. But then \cref{eq:to_prove4} becomes
  \[
    W_u(k_u)W_v(k_v)W_u(\neg k_u)W_v(\neg k_v)C^2 = W_u(k_u)W_v(\neg
    k_v)W_u(\neg k_u)W_v(k_v)C^2
  \]
  which is trivially true.
\end{proof}

@@

\subsection{Changes to ADDMC}

@@

\todo[inline]{This needs rewriting.}
ADDMC multiplies the final answer by $w(u)+w(-u)$ for every $u \in U$ that was
`simplified out' of the ADD for $\phi$ and so is not featured in the $\exists_U$
projection. In our format, for every such $u \in U$, we multiply the answer by
two. Each such $u \in U$ must satisfy two properties:
\begin{itemize}
\item all the probabilities/weights associated with $u$ are equal to $0.5$
  (otherwise the corresponding CPT would depend on $u$),
\item and all other CPTs are independent of $u$ (or they may have a trivial
  dependence, where the probability stays the same if $u$ is replaced with its
  complement).
\end{itemize}
Thus, the CPT that corresponds to $u$ still multiplies every model by $0.5$, but
the number of models being considered by the ADDMC is halved. To correct for
this, we need to perform this multiplication.
